---
title: "sgo - Simple Geographical Operations (with OSGB36)"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  fig.width = 5,
  out.width = "50%"
)
```

<!-- badges: start -->
[![R-CMD-check](https://github.com/clozanoruiz/sgo/workflows/R-CMD-check/badge.svg?branch=main)](https://github.com/clozanoruiz/sgo/actions?workflow=R-CMD-check)
[![coverage](https://codecov.io/gh/clozanoruiz/sgo/branch/main/graph/badge.svg?token=Qd5gkpnxFc)](https://app.codecov.io/gh/clozanoruiz/sgo?branch=main)
[![CRAN](https://www.r-pkg.org/badges/version/sgo)](https://cran.r-project.org/package=sgo)
[![GitHub release (latest SemVer)](https://img.shields.io/github/v/release/clozanoruiz/sgo.svg?sort=semver&label=GitHub)](https://github.com/clozanoruiz/sgo/releases)
[![license](https://img.shields.io/badge/license-BSD%202--Clause-green.svg)](https://opensource.org/licenses/BSD-2-Clause)
<!-- badges: end -->

The goal of `sgo` is to provide a set of methods to transform ETRS89 coordinates to the equivalent
OSGB36 National Grid coordinates, and vice-versa, using the [Ordnance Survey's National Grid Transformation Model OSTN15](https://www.ordnancesurvey.co.uk/blog/2016/09/ostn15-new-geoid-britain/). Additionally, it also has the ability to calculate distances and areas from groups of points defined in any of the coordinate systems supported.

The Coordinate Systems supported by `sgo` are:

* ETRS89: EPSGs 4258, 4937, 4936 and 3035
* WGS84: EPSGs 4326, 4979, 4978 and 3857
* OSGB36: EPSGs 27700, 7405 and 4277

Please note that this package assumes that the Coordinate Reference Systems
(CRS) ETRS89 and WGS84 are the same within the UK, but this shouldn't be a
problem for most civilian use of GPS satellites. If a high-precision
transformation between WGS84 and ETRS89 is required then it is recommended
to use a different package to do the conversion.

According to the Transformations and OSGM15 User Guide, p. 8:
*"...ETRS89 is a precise version of the better known WGS84  system
optimised for use in Europe; however, for most purposes it can be considered
equivalent to WGS84."* and *"For all navigation, mapping, GIS, and engineering
applications within the tectonically stable parts of Europe (including UK and
Ireland), the term ETRS89 should be taken as synonymous with WGS84."*.

## Installation

You can install the released version of `sgo` from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("sgo")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("clozanoruiz/sgo")
```
## Usage

```{r, message = FALSE}
library(sgo)

## sgo_points creation (from lists, matrices or dataframes)
sr <- sgo_points(list(-3.9369, 56.1165), epsg=4326)
sr
ln <- c(-4.22472, -2.09908)
lt <- c(57.47777, 57.14965)
n <- c("Inverness", "Aberdeen")
df <- data.frame(n, ln, lt, stringsAsFactors = FALSE)
locs <- sgo_points(df, coords=c("ln", "lt"), epsg=4326)

library(maps)
map('world', regions=('uk'), xlim=c(-9, 0), ylim=c(54.5, 60.9))
points(x=sr$x, y=sr$y, pch=0, col="green")
points(locs, pch=0, col="red")
text(locs, labels=locs$n, pos=1, cex=0.9)
text(sr, labels="Stirling", pos=1, cex=0.9)

## Convert coordinates to OS National Grid reference
bv <- sgo_points(list(x=247455, y=706338, name="Ben Venue"),
coords=c("x", "y"), epsg=27700)
sgo_bng_ngr(bv)$ngr
# notice the truncating:
sgo_bng_ngr(bv, digits=8)$ngr

## Convert from lon/lat to BNG
lon <- c(-4.25181,-3.18827)
lat <- c(55.86424, 55.95325)
pts <- sgo_points(list(longitude=lon, latitude=lat), epsg=4326)
sgo_lonlat_bng(pts)

## sgo_transform is a wrapper for all the conversions available
# to BNG:
sgo_transform(locs, to=27700)
#
# to OSGB36 (historical):
sgo_transform(locs, to=4277)
#
# to Pseudo-Mercator:
sgo_transform(locs, to=3857)

## Convert OS National Grid references to ETRS89
munros <- data.frame(
  ngr=c("NN 1652471183", "NN 98713 98880",
         "NN9521499879", "NN96214 97180"),
  name=c("Beinn Nibheis", "Beinn Macduibh",
          "Am Bràigh Riabhach", "Càrn an t-Sabhail"))
sgo_transform(sgo_ngr_bng(munros, col="ngr"), to=4258)

## Calculate distances
# Distance in metres from one point to 2 other points
p1 <- sgo_points(list(-3.9369, 56.1165), epsg=4326)
lon <- c(-4.25181,-3.18827)
lat <- c(55.86424, 55.95325)
pts <- sgo_points(list(longitude=lon, latitude=lat), epsg=4326)
p1.to.pts <- sgo_distance(p1, pts, by.element = TRUE)
p1.to.pts

# Perimeter in metres of a polygon defined as a series of ordered points:
lon <- c(-6.43698696, -6.43166843, -6.42706831, -6.42102546,
-6.42248238, -6.42639092, -6.42998435, -6.43321409)
lat <- c(58.21740316, 58.21930597, 58.22014035, 58.22034112,
58.21849188, 58.21853606, 58.21824033, 58.21748949)
pol <- sgo_points(list(lon, lat), epsg=4326)
# Let's create a copy of the polygon with its coordinates shifted one position
# so that we can calculate easily the distance between vertices
coords <- sgo_coordinates(pol)
pol.shift.one <- sgo_points(rbind(coords[-1, ], coords[1, ]), epsg=pol$epsg)
perimeter <- sum(sgo_distance(pol, pol.shift.one, by.element=TRUE))
perimeter

## Area of an ordered set of points
A <- sgo_area(pol)
sprintf("%1.2f", A)
# Interpolate new vertices (here every 10 metres) if more accuracy is needed
A <- sgo_area(pol, interpolate=10)
sprintf("%1.2f", A)
```
## Benchmark

Very simple benchmark in a Intel(R) Core(TM) i7-6700HQ CPU @2.60GHz, and 16 GB of RAM.
See the code of this README file to see the R statements used.
```{r, eval=FALSE, echo=FALSE}
xmin <- -3.5
xmax <- 2.5
ymin <- 53.5
ymax <- 56.5

each_n <- function(n, reps) {

  calc_coords <- function(i) {
    p <- cbind(round(runif(n, xmin, xmax), 8), round(runif(n, ymin, ymax), 8))
    #points(p)
    #map('world', regions=('uk'), xlim=c(xmin, xmax), ylim=c(ymin, ymax))
    #rect(xmin, ymin, xmax, ymax)

    #2D lonlat to bng
    lonlat <- sgo_points(p, epsg=4258)
    start_time <- Sys.time()
    bng <- sgo_transform(lonlat, to=27700)
    t1 <- Sys.time()- start_time

    #2D bng to lonlat
    start_time <- Sys.time()
    bng2ll <- sgo_transform(bng, to=4258)
    t2 <- Sys.time()- start_time

    #3D lonlat to bng
    p.3d <- cbind(p, rep(0, n))
    lonlat.3d <- sgo_points(p.3d, epsg=4937)
    start_time <- Sys.time()
    bng.3d <- sgo_transform(lonlat.3d, to=7405, OD=FALSE)
    t3 <- Sys.time() - start_time

    #3D bng to lonlat
    start_time <- Sys.time()
    bng.3d2ll <- sgo_transform(bng.3d, to=4979, OD=FALSE)
    t4 <- Sys.time() - start_time

    #3D lonlat to bng (OD=TRUE)
    p.3d <- cbind(p, rep(0, n))
    lonlat.3d <- sgo_points(p.3d, epsg=4937)
    start_time <- Sys.time()
    bng.3d <- sgo_transform(lonlat.3d, to=7405, OD=TRUE)
    t5 <- Sys.time() - start_time

    #3D bng to lonlat (OD=TRUE)
    start_time <- Sys.time()
    bng.3d2ll <- sgo_transform(bng.3d, to=4979, OD=TRUE)
    t6 <- Sys.time() - start_time

    out <- c(t6, t4, t2, t5, t3, t1)
  }

  apply(sapply(1:reps, calc_coords), 1, mean)

}

n <- c(1000, 10000, 100000, 1000000, 5000000, 10000000)
reps <- 2
result <- sapply(n, each_n, reps)
```

```{r, echo=FALSE, fig.width = 7, fig.height = 3.5, out.width="100%"}
##### PLOT RESULTS #####
n <- c(1000, 10000, 100000, 1000000, 5000000, 10000000)
result <- matrix(NA,nrow=6, ncol=6)
series <- c("BNG to LL (3D+OD)", "BNG to LL (3D)",  "BNG to LL (2D)",
            "LL to BNG (3D+OD)", "LL to BNG (3D)", "LL to BNG (2D)")
result[1,] <- c(0.006632924, 0.04785049, 0.3512925, 3.412248, 17.286100, 35.01287)
result[2,] <- c(0.005520940, 0.04217410, 0.2820065, 2.825136, 14.126171, 28.38339)
result[3,] <- c(0.008078456, 0.03790104, 0.2718840, 2.567024, 12.890800, 26.27742)
result[4,] <- c(0.004070640, 0.01709044, 0.1517755, 1.505867,  7.350618, 14.95144)
result[5,] <- c(0.003953457, 0.01806092, 0.1187755, 1.205132,  6.302003, 12.52924)
result[6,] <- c(0.004461050, 0.01568353, 0.1435580, 1.070004,  5.667857, 11.35794)

color <- c("#1b9e77", "#d95f02", "#7570b3", "#e7298a", "#66a61e", "#e6ab02")

par(mar=c(5.1, 4.1, 4.1, 11.5), xpd=TRUE)


matplot(n, t(result), type="b", pch = 20, lty=1,  col=color, xaxt = 'n', ylim=c(0,45), xlim=c(1, 10000000),
     main="OSTN15 transformations", xlab=" pairs of coordinates", ylab="seconds")

myTicks = axTicks(1)
axis(1, at = myTicks, labels = formatC(myTicks, format = 'd', big.mark=","))

legend("topright", inset=c(-0.52,0), legend = series,
       lty = c("6a"), pch=20, col = color, box.lty=0, cex=0.8)
```
